### 2025-08-13 actuarial.qmd
20250813 - tweaks

Using the judgment method of prediction, all gathered information is collected and formulated into a prediction in the person's mind.
The person selects, measures, and combines information and produces projections solely according to their experience and judgment.
For instance, a proclaimed "fantasy expert" might use their experience, expertise, and judgment to make a prediction about how each player will perform by using whatever information and data they deem to be important, aggregating all of this information in their mind to make the prediction for each player.
Professional scouts and coaches use judgment when making predictions or selecting players based on their impressions of the players [@DenHartigh2018].
As an example in popular media, in the movie, "Trouble with the Curve", a professional scout makes a judgment about how well a baseball hitter will do in the major leagues from his impressions of the hitter's ability based on the sound of the ball off the player's bat.

@@

Human judgment is naturally subject to errors.
Common [heuristics](#sec-heuristics), [cognitive biases](#sec-cognitiveBiases), and [fallacies](sec-fallacies) are described in @sec-cognitiveBias.
Below, I describe a few errors to which human judgment seems particularly prone.

>>>>

Using the judgment method of prediction, all gathered information is collected and formulated into a prediction in the person's mind.
The person selects, measures, and combines information and produces projections solely according to their experience and judgment.
For instance, a proclaimed "fantasy expert" might use their experience, expertise, and judgment to make a prediction about how each player will perform by using whatever information and data they deem to be important, aggregating all of this information in their mind to make the prediction for each player.
Professional scouts and coaches use judgment when making predictions or selecting players based on their impressions of the players [@DenHartigh2018].
Experts such as scouts and coaches are often former athletes who are presumed to have unique knowledge of their sport.
Because of their expertise, it is possible that they could make observations that a non-expert could not and that could not be captured by statistics [@DenHartigh2018].
As an example in popular media, in the movie, "Trouble with the Curve", a professional scout makes a judgment about how well a baseball hitter will do in the major leagues from his impressions of the hitter's ability based on the sound of the ball off the player's bat.

@@

Human judgment is naturally subject to errors.
Common [heuristics](#sec-heuristics), [cognitive biases](#sec-cognitiveBiases), and [fallacies](sec-fallacies) are described in @sec-cognitiveBias.
Judgments of sports athletes have been shown to be biased as a function of many characteristics of the athlete, including whether they are known to have a positive reputation, and their clothing, body language, and skin color [@DenHartigh2018].
Below, I describe a few errors to which human judgment seems particularly prone.

### 2025-08-13 draft.qmd
20250813 - tweaks

@@

It appears that Running Backs show the fastest dropoff (i.e., steepest negative slope), followed by Tight Ends, Quarterbacks, Wide Receivers, and then defensive positions.
So, identifying the top-scoring Running Back can be of great benefit.

@@

Although Running Backs appear to show the fastest dropoff, Quarterbacks and Wide Receivers have appeared to overtake them in terms of fantasy points.

@@

>>>>

Another approach is to embrace variability and increase the variability of a team's outcomes by [stacking](#sec-stacking) players from the same team [@Lee2022].

@@

Running Backs show the fastest intial dropoff (i.e., steepest negative slope), followed by Tight Ends, Quarterbacks, Wide Receivers, and then defensive positions.
So, identifying the top-scoring Running Back can be of great benefit.
After the top 10 or so players at each position, Quarterbacks show a steeper dropoff than Running Backs, Tight Ends, and Wide Receivers.

@@

Since 2020, although Running Backs show the fastest initial dropoff, Quarterbacks and Wide Receivers have appeared to overtake them in terms of fantasy points.
Quarterbacks continue to show the steepest dropoff after the top 10 or so players at each position.

@@

Key takeaways include that, based on their steep initial dropoff, it can be of great benefit (in terms of a comparative advantage against your opponents) to have a top Running Back or Tight End.
The challenge, of course, is identifying—in advance—which of the players is going to be the top-ranked player at that position.
It also seems important to have a top-10 Quarterback, because Quarterbacks show the steepest dropoff after position rank of ~10.
In general, given their tendency to score lots of points and to show a relatively steep dropoff, it seems crucial to have strong Quarterbacks, Wide Receivers, and Running Backs, with the highest priority likely going to Running Backs and Wide Receivers.

### 2025-08-13 mixed-models.qmd
20250813 - tweaks

Applied to longitudinal fantasy football data, Most simply, players' longitudinal performance data would be modeled as nested data with a two-level model: timepoints nested within players.
However, you could get more complicated.
For instance, you could estimate a three-level model of timepoints nested within players nested within teams.
This would help account for players's productivity benefitting or taking a hit from being on a given team, and it would account for players changing teams.
Alternatively, consistent with @Usami2018, you could estimate a cross-classified model of players' individual observations/measurement occasions belonging to two upper-level units: players and time points (e.g., week 1, week 2, 2025 season, etc.).
This would help account for systematic effects that are specific to a given timepoint (e.g., week or season).

>>>>

Applied to longitudinal fantasy football data, Most simply, players' longitudinal performance data would be modeled as nested data with a two-level model: timepoints nested within players.
However, you could get more complicated.
For instance, you could estimate a three-level model of timepoints nested within players nested within teams.
This would help account for players' productivity benefitting or taking a hit from being on a given team, and it would account for players changing teams.
Alternatively, consistent with @Usami2018, you could estimate a cross-classified model of players' individual observations/measurement occasions belonging to two upper-level units: players and time points (e.g., week 1, week 2, 2025 season, etc.).
This would help account for systematic effects that are specific to a given timepoint (e.g., week or season).

### 2025-08-13 modern-portfolio-theory.qmd
20250813 - tweaks

Stacking involves selecting players on the same team (e.g., multiple players on the Dallas Cowboys).
To help inform the process of stacking, we evaluate the inter-correlation of performance in a given week across players on the same team and on opposing teams.
We determine QB1, RB1, RB2, WR1, WR2, WR3, TE1, and K1 for each combination of team and season based on how many fantasy points a player scored over the entire season.

>>>>

Stacking involves selecting players on the same team (e.g., multiple players on the Dallas Cowboys).
Having two players from the same team increases the variance of the fantasy points scored, for better or for worse, because their performance is linked [i.e., can depend on one another; @Lee2022].
Thus, when the players' respective team performs well, the players may score many points, greatly increasing the likelihood that a fantasy team wins; however, if the players' respective team does not perform well, the players may score few points, thus increasing the likelihood that the fantasy team loses.
It is a higher-risk strategy—but one that can pay off if the players are on a high-scoring offense.
To help inform the process of stacking, we evaluate the inter-correlation of performance in a given week across players on the same team and on opposing teams.
We determine QB1, RB1, RB2, WR1, WR2, WR3, TE1, and K1 for each combination of team and season based on how many fantasy points a player scored over the entire season.

### 2025-08-13 player-evaluation.qmd
20250812 - linked figures
20250812 - unlink figures
20250813 - tweaks

@@

When thinking about the most important domains to assess for evaluating players and predicting their performance, it is important to distinguish between signs and samples [@DenHartigh2018].
*Signs* are indicators of underlying states.
The signs approach to assessment involves assessing processes that may predict performance, where the emphasis is on what the sign indicates about the underlying attribute, rather than on the specific behavior itself.
In football, the signs approach might involve measuring skills using separate tests where players' skills are tested in isolation.
For instance, the NFL Combine takes a signs approach to assessment, in which players have a separate test to assess for speed (e.g., 40-yard dash) and jumping ability (e.g., vertical jump).

>>>>

We demonstrate how to download fantasy football projections in @sec-fantasyFootballProjections.

@@

When thinking about the most important domains to assess for evaluating players and predicting their performance, it is important to distinguish between signs and samples [@DenHartigh2018].
*Signs* are indicators of underlying states.
The signs approach to assessment involves assessing processes that may predict performance, where the emphasis is on what the sign indicates about the underlying attribute, rather than on the specific behavior itself.
In football, the signs approach might involve measuring skills using separate tests where players' skills are tested separately (i.e., in isolation from other skills) in controlled settings.
For instance, the NFL Combine takes a signs approach to assessment, in which players have a separate test to assess for speed (e.g., 40-yard dash) and jumping ability (e.g., vertical jump).
The signs approach to assessment assumes that the various skills assessed (e.g., being fast, being able to jump high, having strong endurance, being motivated) have high [predictive validity](#sec-predictiveValidity) for the target behavior [i.e., future performance in games; @DenHartigh2018].
